---
title: "Regression"
output: html_notebookd
author: "Zachary Canoot & Gray Simpson"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# Regression

Using the Hungary Dataset [Weather in Szeged 2006-2016](https://www.kaggle.com/datasets/budincsevity/szeged-weather)
Found on Kaggle. 

[how linear regression works, with strengths and weaknesses]


First, we divide our data up into training and testing.
```{r}

```


Next, we want to explore our training data. 
We'll look into averages, the head of the data, the range, the correlation, and the summary of it.
```{r}

```


We'll pull up two graphs to get a better idea of what we have to do, now.
```{r}

```


Let's start with a linear regression model with one predictor, and summarize it. 
```{r}

```
[insert explanation of info seen in the summary]

Lets plot the residuals.
```{r}

```
[insert explanation of what the residual plot tells us]


Let's up the complexity, now. We'll build a multiple linear regression model, and see if we can improve the accuracy.
```{r}

```
[short explanation of it]

Now let's go a step even farther. We'll use a combination of predictors, interaction effects, and polynomial regression to see if we can get even more accurate. 
```{r}

```


[long explanation talking about which model was the best, comparing the results, and explaining why we liked which model the best]


Using the three models, we will predict and evaluate using the metric correlation and MSE. 
```{r}

```
[insert explanation, comparinig the results and indicating why we think they happened]



